Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-12-30 manuscript.Rmd
Add revisions to permutation sections

author:
  - name          : "Megan Chiovaro"
    affiliation   : "1, 2"
    corresponding : yes    # Define only one corresponding author
    address       : "406 Babbidge Road, Unit 1020, Storrs, CT 06269"
    email         : "megan.chiovaro@uconn.edu"
  - name          : "Leah C. Windsor"
    affiliation   : "3,*"
  - name          : "Alistair Windsor"
    affiliation   : "4"
  - name          : "Alexandra Paxton"
    affiliation   : "1,2,*"

@@

The tweets were first sorted into ascending order by timestamp. They were then
grouped into successive sets of five tweets. Groups of five were chosen to allow
for manageability and efficiency of the analyses. Linguistic cohesion was
quantified for all possible pairs of tweets in the group of five, resulting in
10 pairs per group of tweets; all 10 cohesion values for the group were averaged
for a single group cohesion value. The cohesion value for each day was
quantified as the average cohesion values for all groups in that day. Each
pair’s cohesion value was calculated in three parts:

@@

#### Permutation Test
Unlike some nonlinear analyses (e.g., fractal analyses), CRQA metrics do not
have inherently meaningful values. As a result, CRQA is best used as a relative
metric---that is, comparing the CRQA values across conditions or subsets of the
data. Typically, meaningful differences between values are determined with
traditional inferential statistics. While this may not be a problem for
experimentally derived datasets, it can pose a barrier to researchers using
smaller datasets or conducting data-informed case studies.

For CRQA, this can be addressed using *permutation tests*. Similar to more
established methods of statistical significance in CRQA research [which create
surrogate time series; e.g., @paxton2017interpersonal], the permutation test
approach to significance testing allows the researcher to use a time series as
its own baseline by testing whether the *structure* of the unfolding events of
the time series---rather than the *raw frequency* of the events within the time
series---cohere together more than would be expected by chance. A permutation
test is similar to bootstrapping, except that the sampling of observations is
done without replacement. In this approach, we simply create a large number of
permutations of the original time series (i.e., breaking up the dependency
across time points but preserving the frequencies of the events), conduct CRQA
for each of those permutations, and then examine whether the real time series'
CRQA metrics exceed those that we would expect to see by chance. The proportion
of times that the real time series' values exceeds the surrogate time series'
values can be interpreted as the alpha criterion to determine significance. For
more information on permutation tests, see @good2013permutation.

Using the real deciled time series of social cohesion, count of all events,
count of positive events, and count of negative events, we conducted a series of
permutation tests (one for the comparison between social cohesion and each type
of event count) to determine whether the observed measure is statistically different from chance for each CRQA
metric. Using the `sample` function in base R [@R-base], $1000$ permuted time
series were created for each of the real time series. CRQA was then run $1000$
times for social cohesion and count of all events, social cohesion and count of
positive events, and social cohesion and count of negative events using the
permuted time series. This resulted in $1000$ results for each of the CRQA
metrics for each combination of time series. Statistical significance from
chance was set at the 95th and 99th percentiles of the metrics from the permuted
CRQ analyses. Given that a permutation test is similar to bootstrapping but does
not include replacement, all real recurrent points were included in each CRQA
analysis, leaving RR constant across all permutations. Thus, RR is not included
in the inferential tests.

@@

#### Parameters
In addition to the CRQA parameters, two other parameters are required when
conducting WCRQA. First, the *window size* will determine how many observations
are used in each CRQA analysis. We conducted WCRQA with a window size of $14$ to
investigate the dynamics over two weeks and allow for the capture of more
structure and thus a more meaningful interpretation of DET. A window size
smaller than 14 would limit the possibilities for shared trajectories and make
determinism uninterpretable, and window sizes larger than 14 would increasingly
overshadow the fine-grain dynamics that we want to investigate.

@@

#### Permutation Test
To visualize the differences in the WCRQA metrics from those obtained by chance,
a permutation test was done to set upper and lower bounds for the 95th and 99th
percentiles. Using a sample size of 14, $1000$ permuted time series were
generated without replacement for each real time series. The upper and lower
95th and 99th percentiles were set based on the distribution of the permuted
WCRQA metrics.

@@

### Syria as Source and Target
For count of all events and social cohesion, the WCRQA plot for RR looked
largely the same as that of the target-only events, but DET showed a new early
peak in above-chance values. This suggests that the overall number of events
were more consistent from day to day when considering source events, as might be
expected given that individuals can only participate in a limited number of
events each day. Days that previously had a lower than average number of target
events may now have had a more near-average number of source and target events
because the reduced participation in target events allows for increased
participation and initiation of source events.

@@

## Conclusion
Considering the pervasiveness of social media, it is unsurprising that social
activists and protesters have utilized these platforms to support real-world
mobilization. Here, we quantitatively explored the connection between online
social cohesion and real-world events. In addition to introducing political
scientists to a family of powerful nonlinear dynamics analyses, our results have
serious implications for monitoring global movements. While state-run media in
regions of authoritarian rule are often biased and serve as propaganda for the
leader and central government, the self-organized nature of Twitter provides a
new avenue to gain insight into truths about uprisings and social mobilization.
By conceptualizing Twitter and real-world events as two elements of an
inextricably linked dynamical system, the levels of social cohesion within
social media platforms could be monitored for fluctuations indicating shifts in
power and peace. Social media mobilization may thus help alert the world to
citizens' actions and reactions that are otherwise suppressed by government
regimes, providing insight into events that may not otherwise be broadcast and
allowing for faster intervention.

>>>>

author:
  - name          : "Megan Chiovaro"
    affiliation   : "1, 2"
    corresponding : yes    # Define only one corresponding author
    address       : "406 Babbidge Road, Unit 1020, Storrs, CT 06269"
    email         : "megan.chiovaro@uconn.edu"
  - name          : "Leah C. Windsor"
    affiliation   : "3,*"
  - name          : "Alistair Windsor"
    affiliation   : "3,4"
  - name          : "Alexandra Paxton"
    affiliation   : "1,2,*"

@@

The tweets were first sorted into ascending order by timestamp. They were then
grouped into successive sets of five tweets. Groups of five were chosen to allow
for manageability and efficiency of the analyses. Linguistic cohesion was
quantified for all possible pairs of tweets in the group of five, resulting in
10 pairs per group of tweets; all 10 cohesion values for the group were averaged
for a single group cohesion value. The cohesion value for each day was computed
as the average cohesion values for all groups whose first tweet was on that day.
Each pair’s cohesion value was calculated in three parts:

@@

#### Approximate Permutation Test
Unlike some nonlinear analyses (e.g., fractal analyses), CRQA metrics do not
have inherently meaningful values. As a result, CRQA is best used as a relative
metric—that is, comparing the CRQA values across conditions or subsets of the
data. Typically, meaningful differences between values are determined with
traditional inferential statistics. While this may not be a problem for
experimentally derived datasets, it can pose a barrier to researchers using
smaller datasets or conducting data-informed case studies.

For CRQA, this can be addressed using *approximate permutation tests*. Similar
to more established methods of statistical significance in CRQA research (which
create surrogate time series; e.g., @paxton2017interpersonal), the permutation
test approach to significance testing allows the researcher to use a time series
as its own baseline by testing whether the *structure* of the unfolding events
of the time series—rather than the *raw frequency* of the events within the time
series—cohere together more than would be expected by chance. Permutation tests
are used to identify null distributions for test statistics when the null
hypothesis is that the group assignment does not matter, and they can be used
when traditional parametric tests may not be appropriate [e.g.,
@collingridge2013primer;
@boring2016student]. The test is similar to bootstrapping, except that the
sampling of observations is done without replacement.

To quantify the differences in the CRQA metrics from those that would be
expected by chance, we conducted approximate permutation tests for significance
with upper and lower bounds for the 95th and 99th percentiles (comparable to
traditional alpha criteria of .05 and .01, respectively). In this approach, we
simply create a large number of permutations of the original time series (i.e.,
breaking up the dependency across time points but preserving the frequencies of
the events), conduct CRQA for each of those permutations, and then examine
whether the real time series’ CRQA metrics exceed those that we would expect to
see by chance. The proportion of times that the real time series’ values exceeds
the surrogate time series’ values can be interpreted as the alpha criterion to
determine significance. For more information on permutation tests, see
@good2013permutation.

Using the real deciled time series of social cohesion, count of all events,
count of positive events, and count of negative events, we conducted a series of
permutation tests (one for the comparison between social cohesion and each type
of event count) to determine whether the observed measure is statistically
different from chance for each CRQA metric. Using the `sample` function in base
R [@R-base], $1000$ permuted time series were created for each of the real time
series. CRQA was then run for each permuted time series combination for social
cohesion and count of all events, social cohesion and count of positive events,
and social cohesion and count of negative events using the permuted time series.
This resulted in $1000$ results for each of the CRQA metrics for each
combination of time series. Statistical significance from chance was set at the
95th and 99th percentiles of the metrics from the permuted CRQ analyses. Given
that a permutation test is similar to bootstrapping but does not include
replacement, all real recurrent points were included in each CRQA analysis,
leaving RR constant across all permutations. Thus, RR is not included in the
inferential tests.

@@

#### Parameters
In addition to the CRQA parameters, two other parameters are required when
conducting WCRQA. First, the *window size* will determine how many observations
are used in each CRQA analysis. We conducted WCRQA with a window size of $14$ to
investigate the dynamics over two weeks and allow for the capture of more
structure and thus a more meaningful interpretation of DET. We first attempted
to use a window size of $7$ to investigate the dynamics at the week scale, but
found that window sizes smaller than $14$ limited the possibilities for shared
trajectories and made determinism uninterpretable. In addition, window sizes
larger than $14$ increasingly overshadowed the fine-grain dynamics that we wanted
to investigate. Thus we decided to use a window size of $14$ to investigate the
dynamics at the two week scale.

@@

#### Approximate Permutation Test
To visualize the differences in the WCRQA metrics from those obtained by chance,
a permutation test was done. $1000$ permuted time series were generated (without
replacement) of length $14$, (in accordance with the 14-day windows of real data
for our WCRQA analyses), for each real time series; WCRQA was conducted on each
combination of cohesion and event permuted time series. The upper and lower 95th
and 99th percentiles for significance were set based on the distribution of the
permuted WCRQA metrics.

@@

### Syria as Source and Target
For count of all events and social cohesion, the WCRQA plot for RR looked
largely the same as that of the target-only events, but DET showed a new early
peak in above-chance values. This suggests that the overall number of events
were more consistent from day to day when considering both source and target
events, as compared to target-only events. In other words, days that previously
had a more variable number of target events may now have had a more stable
number of source and target events.

@@

## Conclusion
Considering the pervasiveness of social media, it is unsurprising that social
activists and protesters have utilized these platforms to support real-world
mobilization. Here, we quantitatively explored the connection between online
social cohesion and real-world events. In addition to introducing political
scientists to a family of powerful nonlinear dynamics analyses, our results have
serious implications for monitoring global movements. While state-run media in
regions of authoritarian rule are often biased and serve as propaganda for the
leader and central government, the self-organized nature of Twitter provides a
new avenue to gain insight into truths about uprisings and social mobilization.
By conceptualizing Twitter and real-world events as two elements of an
inextricably linked dynamical system, the levels of social cohesion within
social media platforms could be monitored for fluctuations indicating shifts in
power and peace.
